---
title: "My Project Data"
author: Camila Sanmiguel
date: 10/01/19
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)

# --------------------------------------------------

# I have read in the data from different spreadsheets
# containing data on Border Patrol apprehensions.
# I will probably be replacing these with a more 
# comprehensive csv, which I'm still looking for online,
# that hopefully contains data for more years.
bp17 <- read_xlsx("bp-apprehensions-17.xlsx")
bp18 <- read_xlsx("bp-apprehensions-18.xlsx")
bp19 <- read_xlsx("bp-apprehensions-19.xlsx")
bp_since <- read.csv("bp-monthly-apprehensions-since-2000.csv")

# --------------------------------------------------

```
# Plan
This is my current plan and collection of data that I'll be springboarding my project off of! 

<span style="color:firebrick">
**I want to examine Mexican cartel violence and its relationship to illegal immigration -- do spikes in incidence of cartel violence galvanize efforts to leave Mexico? How closely do they correlate if at all?**
<br>
<br>
Do spikes in cartel violence cause spikes in illegal immigration to the U.S.? 
<br>
<br>
I want to show progression over the years of large occurrences of violence beside the progression of border apprehensions.</span>

# Data
As of now, I will be needing two data sources for this: the first is from the CBP.
<br>
2019: https://www.cbp.gov/newsroom/stats/sw-border-migration/usbp-sw-border-apprehensions
<br>
2017: https://www.cbp.gov/newsroom/stats/usbp-sw-border-apprehensions-fy2017# 


<br>


The second source of data is from various Mexican government websites and sources. 
<br>
The main one is publicly available from the Mexican government's website.


There are a few other sources on https://elcri.men/en/datos.html that I'm exploring still!

```{r, include=FALSE, echo=FALSE}

# --------------------------------------------------
# Below I read in the data from the government website's 
# reports on homicide and cartel activity.

read.csv("CNSP-Delitos-2019_ago19.csv")

# --------------------------------------------------
```
<br>
I want to include a source that more definitively addresses specific cartels, but I'm not yet sure how to locate research on this - still searching! For now, the aforementioned are my sources.


